A Constructive Ordered Predicate Logic for Category Theory

### 2022-10-11 paper/paper.tex
get the bib compiling
pass over the intro (not any shorter uh oh)

\usepackage{tikz-cd} \usepackage{amssymb} \usepackage{amsmath}
%% \usepackage{amsthm}
\usepackage{mathtools}
\usepackage{stmaryrd}
\usepackage{mathpartir}
\usepackage{bbm}
\usepackage{cleveref}
\usepackage{natbib}

@@

@@

Category theory is a branch of mathematics that gives a systematic
theory of typed, ``higher-dimensional'' algebraic structures.
%
Originally developed for applications to homological algebra, it was
quickly discovered that (among many other applications) these
categorical structures were common in logic and computer science, and
that formal systems like logics, type theories and programming
languages can be soundly and completely modeled
by notions of structured categories\cite{lawvere,goguen,lambek,moggi}, a
phenomenon sometimes referred to as the Curry-Howard-Lambek
nncorrespondence \cite{curry-howard-lambek}.
%
Examples include the simply typed lambda calculus \cite{lambekscott},
computational lambda calculus \cite{moggi}, linear logic \cite{girard}
and dependent type theory\cite{cartmell1986}.
%
These seminal works have led to a fruitful field of research known as
categorical logic, where new type theories are designed based directly
on category-theoretic semantics.
%
The slogan is that the syntax of a type theory should present an
initial object in its category of models, a category-theoretic
reformulation and generalization of the logical notion of soundness
and completeness.

% TODO: do we have citations for this?
While this research program has been quite successful,
category-theoretic techniques in computer science are sometimes
criticized for being overly complex, and there is difficulty in
acquisition of competence in category theory by computer scientists
who do not have a strong background in pure mathemantics.
%
While much of this can be explained by lack of, or lack of awareness
of, accessible material for computer scientists, in some ways category
theory is objectively complex.
%
In a traditional set-theoretic formulation, notions such as adjoint
functors and Kan extensions produce a proliferation of ``naturality''
and ``functoriality'' side-conditions that must be discharged.
%
This is compounded when we move from ordinary category theory to
enriched and internal category theory, where constructions must be
additionally proven continuous, monotone, etc, in addition to natural
or functorial.
%
These generalizations have long been recognized as being important in
programming language applications: domain-, metric- and
step-index-enriched categories have been used to model general
recursion and internal categories have been used to model
parametricity and gradual
typing\cite{order-enriched-cats,topos-of-trees,reflexive-graphs,double-cats-gradual-typing},
to name a few.

% TODO: Yoneda example
Novices in category theory are often overwhelmed by the number of
naturality and functoriality side conditions that abound in otherwise
simple proofs in category theory. For instance consider the proof of the
Yoneda lemma, often the first ``big'' proof in an introductory category
theory text. The Yoneda lemma states that for any category $C$, functor
$P : C^{o} \to \Set$ (where $C^{o}$ is the ``opposite'' of $C$, i.e. 
$C$ with the direction of morphisms reversed) and object $A \in C$, the
set of natural transformations $C(-,A) \to P$ is in bijection with the
elements $P(A)$. Furthermore, this bijection is natural in both $A$ and
$P$. In the authors' experience teaching category theory the sheer
amount of naturality side conditions in this result can be very
intimidating to learners, while experienced practitioners will observe
that these naturality conditions are tedious to carry out but
routine. In the type theory we introduce in this paper, none of these
naturality conditions need to be explicitly proven. Instead, all terms
expressible in the type theory are automatically natural by
construction.

Fortunately, the tools of category theory itself can be employed to
simplify these constructions, specifically the tools of \emph{higher}
category theory.
%
Consider the analogous situation in differential calculus. When an
adept analyst writes down a function, they do not expand out the
$\epsilon\mathord{-}\delta$ definition of continuity and prove that it is
continuous. Rather, they are aware of certain \emph{syntactic
principles}, most importantly that composition of continuous functions
is continuous and that previously defined functions are known to be
continuous.
%
Similar principles apply to category theory itself: a composition of
functors is functorial, a composition of natural transformations is
natural, etc.
%
In the case of category theory, these principles can be formalized
using algebraic structures such as 2-categories, bicategories,
(virtual) double categories and pro-arrow equipments\cite{}.
%
This approach has the advantage of avoiding the need to prove
the side conditions of naturality or enrichment.
%
A downside is that these algebraic structures are quite complicated,
and practitioners typically employ either an algebraic combinator
syntax or a 2-dimensional diagrammatic language that can be quite
beautiful and elegant, but is also seems very far removed from the
traditional formulation of category theory in terms of sets and
functions.

In this work, we apply the technique of categorical logic to these
algebraic structures to design a formal logic, which we call
\emph{virtual equipment type theory}, that can be soundly interpreted
in virtual equipments.
%
In the ``standard model'' of \vett{}, the sorts are interpreted as
categories, functions as functors, relations as profunctors and
implications of relations as profunctor homomorphisms, a generalization
of natural transformations.
%
The type theoretic syntax allows us to write all of our constructions
as if they were ordinary functors and profunctors in set-based
category theory.
%
\vett also can be modeled in an arbitrary virtual equipments, and so the
constructions are valid in enriched and internal category theory as
well.
%%
%% DRL: I know what you mean, but I think conflating set theory and
%% type theory will put some people off... 
%% 
%% This is analogous to the situation in intuitionistic type theory,
%% where the syntax is similar to that of informal set theory, but by
%% restricting to intuitionistic reasoning, the constructions can all be
%% interpreted in models where types are spaces and terms are continuous
%% functions ``for free''.
%
The syntax provides an alternative to algebraic and
string-diagrammatic syntaxes for working with virtual equipments,
similar to how the lambda calculus provides an alternative to
categorical combinators and string diagram calculi for cartesian
closed categories.

\vett{} is an \emph{ordered}, \emph{constructive} variant of
predicate logic over a unary type theory.
%
It is \emph{ordered} in that our notion of predicate is that of a
binary relation $R(\alpha;\beta)$ where the $\alpha$ is thought of as
a contravariant dependence and the $\beta$ has a covariant dependence,
and the rules of the logic distinguish between these two variances.
%
Proofs in the predicate logic $R_1, R_2,\ldots \vdash S$ are
restricted as in ordered linear logic \cite{lambek-calculus, ordered-linear-logic} in
that each input must be used exactly once and in the order they were
declared.
%
This order discipline ensures that every proof in the logic can be
interpreted as a \emph{natural} transformation.
%
The logic is \emph{constructive} in that we care about the identity of
proofs: they are natural transformations and there may be more than
one natural transformation of the same type.
%
If we ignore the identity of proofs, ordered predicate logic can be
viewed as a logic of preordered sets rather than categories.

The restriction to unary type theory and binary relations is fairly
strict: we cannot discuss important concepts such as functor
categories, opposite categories and arbitrary (co)-ends of
profunctors. However, the restriction is natural in that it
corresponds to a well-understood notion of model (virtual equipments)
and so we leave these extensions to future work.
%
It is well known that a great deal of fundamental results and
constructions in category theory such as Yoneda and co-Yoneda lemmas,
adjoint functors and Kan extensions are already expressible in virtual
equipments, for instance virtual equipments are used for this purpose
in a text on $\infty$-category theory to avoid explicit manipulation
of model categories \cite{elements-of-oo-category-theory}.
%
We demonstrate some of these standard constructions
in \ref{sec:synthetic} to show that \vett{} gives a syntax in which
these formal constructions look like standard set-theoretic arguments.

However, we can work around these unary/binary restrictions to some
extent by viewing the type theory as a domain-specific language embedded
in a metalanguage.  For example, while we cannot talk about functor
categories inside the ordered part of the type theory, we can state a
theorem that quantifies over functors using the meta-language's
universal quantifier.  To more easily formalize these metatheorems, we
add a third layer to the type theory: an ordinary dependent type theory
in the style of Martin-L\"of type theory. Then all of our ordered
predicate logic judgments will also be indexed by a context from this
dependent type theory, and the type theory includes types for
categories, functors, profunctors and natural transformations that allow
us to formalize theorems the logic is too restrictive to encode. This is
analogous to the notion of a \emph{2-level type
  theory}\cite{voevodsky12,altenkirchXX}: we add a dependent type theory
for reasoning about the terms of our virtual equipment type theory.
The dependent type theory layer also has an additional purpose: since
we can quantify over the structures in \vett{}, we can additionally
encode \emph{signatures} for \vett{} as simply contexts of the
dependent type theory. This allows for a clean formalization of many
categorical structures such as monoidal categories, or strong monads.

While we emphasize the applications to enriched and internal category
theory in this work, there is potential for more direct application to
programming language semantics.
%
Ordinary predicate logic has been used to develop logics in which
logical relations proofs can be developed such as Abadi-Plotkin logic
for parametricity and LSLR and Iris for step-indexed logical relations
proofs \cite{abadi-plotkin,lslr,iris}.
%
We conjecture that \vett{} might similarly serve as the foundation for
a logic of \emph{ordered} structures, which abound in applications:
rewriting and approximation relations can both be modeled as orderings
and logical relations involving these structures are proven to respect
orderings: operational logical relations must be downward-closed and
approximation relations should satisfy transitivity.
%
Just as LSLR and Iris release the user from the syntactic burden of
explicit step-indexing, \vett{} may be used to release the user from
the syntactic burden of proving downward-closure or transitivity
side-conditions.

@@

\item In \cref{sec:examples} we show how to use our syntax for
  synthetic category theory. Much of this is just rendering well-known
  2-categorical/double categorical constructions in our type theoretic
  syntax to demonstrate the ``naive''/synthetic character of proofs
  and constructions in SCT.

%% \item In \cref{sec:synthetic} we show how common theorems of category
%%   theory such as the (co)-Yoneda lemma, equivalence of different
%%   formulations of adjunctions, and preservation of limits can all be
%%   cleanly reproduced in SCT.
%% \item In \cref{sec:internal}, we show how we can use SCT to define
%%   \emph{internal categories}, which allows us to prove additional
%%   theorems that are not formalizable as synthetic theorems.

\item In \cref{sec:semantics}, we develop some basic model theory for
  \vett{}, including a sound and complete notion of categorical model
  and sound interpretation in virtual equipments modeling ordinary,
  enriched and internal category theory.

@@

\begin{definition}[Virtual Equipment]
  A virtual equipment $\mathcal V$ consists of
  \begin{enumerate}
  \item A category $V_o$ of ``objects and vertical arrows''
  \item A chosen subset of the objects $V_s \subseteq V_o$ of ``small objects''
  \item A set ${\mathcal V}_h$ of ``horizontal arrows'' with source
    and target functions $s,t : {\mathcal V}_h \to {\mathcal V_o}^2$
  \item Sets of 2-cells of the following form:
% https://q.uiver.app/?q=WzAsNixbMCwwLCJDXzAiXSxbMCwyLCJEXzAiXSxbMiwwLCJDX3tufSJdLFsxLDAsIlxcY2RvdHMiXSxbMiwyLCJEXzEiXSxbMSwxLCJcXHBoaSJdLFszLDIsIlJfbiIsMCx7InN0eWxlIjp7ImJvZHkiOnsibmFtZSI6ImJhcnJlZCJ9fX1dLFswLDEsIkYiLDJdLFsyLDQsIkciXSxbMSw0LCJTIiwyLHsic3R5bGUiOnsiYm9keSI6eyJuYW1lIjoiYmFycmVkIn19fV0sWzAsMywiUl8wIiwwLHsic3R5bGUiOnsiYm9keSI6eyJuYW1lIjoiYmFycmVkIn19fV1d
\[\begin{tikzcd}
	{C_0} & \cdots & {C_{n}} \\
	& \phi \\
	{D_0} && {D_1}
	\arrow["{R_n}", "\shortmid"{marking}, from=1-2, to=1-3]
	\arrow["f"', from=1-1, to=3-1]
	\arrow["g", from=1-3, to=3-3]
	\arrow["S"', "\shortmid"{marking}, from=3-1, to=3-3]
	\arrow["{R_0}", "\shortmid"{marking}, from=1-1, to=1-2]
\end{tikzcd}\]
    With appropriate ``multi-categorical'' notions of identity and
    composition.

    We say that the 2-cell $\phi$ has $S$ as codomain, the sequence
    $R_0 \ldots R_n$ as domain and call $f$ and $g$ the left and right
    ``frames'', or that $\phi$ is framed by $f$ and $g$.
  \end{enumerate}
  We say a virtual double category has \emph{split restrictions} when
  for any horizontal arrow $R : C \pto D$ and vertical arrows $f : C'
  \to C$ and $g : D' \to D$ there is a chosen horizontal arrow $R(f,g)
  : C' \pto D'$ with a cartesian 2-cell to $R$ framed by $f,g$ and
  these chosen cartesian lifts are functorial in $f,g$.

@@

%% %% Please use bibtex, 
\subsubsection*{Acknowledgements}
Grants stuff...? The authors would like to thank David Jaz Myers, Emily Riehl, Mike
Shulman, Dominic Verity,...for helpful feedback on this work

\bibliography{cats}

>>>>

\usepackage{tikz-cd} \usepackage{amssymb} \usepackage{amsmath}
%% \usepackage{amsthm}
\usepackage{mathtools}
\usepackage{stmaryrd}
\usepackage{mathpartir}
\usepackage{bbm}
\usepackage{cleveref}

@@

\newcommand{\citet}[1]{\cite{#1}}
\newcommand{\citep}[1]{(\cite{#1})}

@@

Category theory is a branch of mathematics that studies
higher-dimensional typed algebraic structures .
%
Originally developed for applications to homological algebra, it was
quickly discovered that (among many other applications) these
categorical structures were common in logic and computer science, and
formal systems like logics, type theories and programming languages
typically have sound and complete models given by notions of
structured categories\cite{lawvere,goguen,lambek,moggi}, a phenomenon
sometimes referred to as the Curry-Howard-Lambek nncorrespondence
\cite{curry-howard-lambek}.
%
Examples include the simply typed lambda calculus \cite{lambekscott},
computational lambda calculus \cite{moggi}, linear logic \cite{girard}
and dependent type theory\cite{cartmell1986}.
%
These seminal works have led to a fruitful field of research known as
categorical logic, where new type theories are designed based directly
on category-theoretic semantics.
%
The slogan is that the syntax of a type theory should present an
initial object in its category of models, a category-theoretic
reformulation and generalization of the logical notion of soundness
and completeness.

% TODO: do we have citations for this?
While this research program has been quite successful,
category-theoretic techniques in computer science are sometimes
criticized for being overly complex, and there is difficulty in
acquisition of competence in category theory by computer scientists
who do not have a strong background in pure mathemantics.
%
While much of this can be explained by lack of, or lack of awareness
of, accessible material for computer scientists, in some ways category
theory is objectively complex.
%
In a traditional set-theoretic formulation, notions such as adjoint
functors and limits produce a proliferation of ``naturality'' and
``functoriality'' side-conditions that must be discharged.
%
This is compounded when we move from ordinary category theory to
enriched and internal category theory, where constructions must be
additionally proven continuous, monotone, etc, in addition to natural
or functorial.
%
These generalizations have long been recognized as being important in
programming language applications: domain-, metric- and
step-index-enriched categories have been used to model recursive
programming languages and internal categories have been used to model
parametricity and gradual
typing\cite{order-enriched-cats,topos-of-trees,reflexive-graphs,double-cats-gradual-typing}.

For instance, when constructing an adjoint pair of functors between
two categories, a na\"ive approach would define all of the data of the
action on objects, action on arrows, prove the functoriality of such
actions, as well as construct two families of transformations, prove
they are natural and then finally proving a pair of equalities
relating compositions of natural transformations.
%
Furthermore, when moving into the enriched or internal setting, the
proofs can have further continuity, monotonicity, or other
side-conditions.
%
Carrying out the proofs explicitly is quite tedious and many newcomers
to are left with the impression that category theory is full of long,
but ultimately trivial constructions.

Fortunately, the tools of category theory itself can be employed to
simplify these constructions, specifically the tools of \emph{higher}
category theory.
%
Consider the analogous situation in differential calculus. When an
adept analyst writes down a function, they do not expand out the
$\epsilon\mathord{-}\delta$ definition of continuity for a function
and proceed from first principles. Rather, they are aware of certain
\emph{syntactic principles} for defining functions that are continuous
by construction, most importantly that composition of continuous
functions is continuous and that previously defined functions are
known to be continuous.
%
Similar principles apply to category theory itself: functors and
natural transformations are closed under composition and whiskering
operations, and experienced category theorists rely on these syntactic
principles to eliminate the tedium of explicit proofs.
%
In the case of category theory, these principles can be formalized
using algebraic structures such as 2-categories, bicategories,
(virtual) double categories and pro-arrow equipments\cite{TODO}, an
approach known as \emph{formal category theory}.
%
In these structures, rather than defining notions of category, functor
and natural transformation from first principles, they are axiomatized
in a manner similar to how a category axiomatizes a notion of space
and homomorphism.
%
One benefit is that the formal approaches also apply to enriched and
internal settings, so \emph{formal} category theory proofs apply to
these settings.
%
A downside is that these algebraic structures are quite complicated,
and practitioners typically employ either an algebraic combinator
syntax or a 2-dimensional diagrammatic language that can be quite
beautiful and elegant, but is also seems very far removed from the
traditional formulation of category theory in terms of sets and
functions.

In this work, we apply the techniques of categorical logic to define a
more familiar logical syntax for carrying out constructions and proofs
in formal category theory.
%
We call the resulting theory \emph{virtual equipment type theory}
(VETT) as (hyperdoctrines of) virtual equipments provide a sound and
complete notion of model for the theory.
%
In VETT, functors and natural transformations are defined using
ordinary term syntax, using familiar notations and $\beta\eta$
reasoning principles for $\lambda$-functions, bound variables, tuples,
etc.
%
By adhering to a \emph{syntactic discipline}, the logic guarantees
that all functor terms are automatically functorial, all and natural
transformation terms are natural.
%
More specifically, the syntax for transformations is a kind of
\emph{indexed, ordered linear} $lambda$ calculus, where the indexing
ensures that transformations are correctly natural and the ordering
and linearity ensure that the proofs are valid in a large class of
enriched and internal categories: for instance enrichment in a
non-symmetric monoidal category.
%
VETT provides an alternative alternative to algebraic and
string-diagram syntaxes for working with virtual equipments,
similar to how the lambda calculus provides an alternative to
categorical combinators and string diagram calculi for cartesian
closed categories.

In this work, we apply the technique of categorical logic to these
algebraic structures to design a formal logic, which we call
\emph{virtual equipment type theory}, that can be soundly interpreted
in virtual equipments.
%
In the ``standard model'' of \vett{}, the sorts are interpreted as
categories, functions as functors, relations as profunctors and
implications of relations as profunctor homomorphisms, a generalization
of natural transformations.
%
The type theoretic syntax allows us to write all of our constructions
as if they were ordinary functors and profunctors in set-based
category theory.
%
\vett also can be modeled in an arbitrary virtual equipments, and so the
constructions are valid in enriched and internal category theory as
well.
%%
%% DRL: I know what you mean, but I think conflating set theory and
%% type theory will put some people off... 
%% 
%% This is analogous to the situation in intuitionistic type theory,
%% where the syntax is similar to that of informal set theory, but by
%% restricting to intuitionistic reasoning, the constructions can all be
%% interpreted in models where types are spaces and terms are continuous
%% functions ``for free''.
%

The syntax of \vett{} is a kind of indexed, ordered linear,
proof-relevant variant of predicate logic over a unary type theory.
%
That is just as a predicate logic has a notion of type, term,
relation and implication, \vett{} is based on four analogous category
theoretic concepts: categories, functors, profunctors and natural
transformations of profunctors.
%
So categories are simply treated like types, and functors are simply
terms whose type is a category and whose free variables have
categories as types as well. The functor syntax is a unary type
theory, meaning functor terms always have precisely one free variable
representing the input object.
%
The analogue of a relation is a \emph{profunctor} (defined below),
which are written as a kind of set with free category variables. In
keeping with our models, like the restriction to unary functors, we
restrict to profunctors with two free variables.
%
The logic is proof-relevant in that the implications of relations are
generalized to natural transformations of profunctors, and we use a
$\lambda$-calculus notation to describe these ``proof terms''.
%
We can strenghten this analogy into a formal construction: any
construction in \vett{} can be erased to a corresponding construction
or proof in predicate logic, as sets, functions and relations define a
(somewhat degenerate) virtual equipment.
%
Our syntax highlights this analogy in that some of our constructions
are generalizations of set comprehension or power-sets and we use
similar notation.

The restriction to unary type theory and binary relations is fairly
strict: we cannot discuss important concepts such as functor
categories or opposite categories. However, the restriction is natural
in that it corresponds to a well-understood notion of model (virtual
equipments).
%
It is well known that a great deal of fundamental results and
constructions in category theory, and they have been used for this
purpose in the formal category theory literature
\cite{elements-of-oo-category-theory, enriched-indexed-categories}.
%
However, we can work around these unary/binary restrictions to some
extent by viewing the type theory as a domain-specific language
embedded in a metalanguage.  For example, while we cannot talk about
functor categories inside the ordered part of the type theory, we can
state a theorem that quantifies over functors using the
meta-language's universal quantifier.  To more easily formalize these
metatheorems, we add a third layer to the type theory: an extensional
dependent type theory in the style of Martin-L\"of type theory. Then
all of our ordered predicate logic judgments will also be indexed by a
context from this dependent type theory, and the type theory includes
universe types for categories, functors, profunctors and natural
transformations that allow us to formalize theorems the logic is too
restrictive to encode, analogous to the use of 2-level type theory in
homotopy type theory \cite{voevodsky12,altenkirchXX}.

%% This is analogous to the notion of a
%% \emph{2-level type theory}: we add a
%% dependent type theory for reasoning about the terms of our virtual
%% equipment type theory.  The dependent type theory layer also has an
%% additional purpose: since we can quantify over the structures in
%% \vett{}, we can additionally encode \emph{signatures} for \vett{} as
%% simply contexts of the dependent type theory. This allows for a clean
%% formalization of many categorical structures such as monoidal
%% categories, or strong monads.

While we emphasize the applications to enriched and internal category
theory in this work, there is potential for more direct application to
programming language semantics.
%
Ordinary predicate logic has been used to develop logics in which
logical relations proofs can be developed such as Abadi-Plotkin logic
for parametricity and LSLR and Iris for step-indexed logical relations
proofs \cite{abadi-plotkin,lslr,iris}.
%
We conjecture that \vett{} might similarly serve as the foundation for
a logic of \emph{ordered} structures, which abound in applications:
rewriting and approximation relations can both be modeled as orderings
and logical relations involving these structures are proven to respect
orderings: operational logical relations must be downward-closed and
approximation relations should satisfy transitivity.
%
Just as LSLR and Iris release the user from the syntactic burden of
explicit step-indexing, \vett{} may be used to release the user from
the syntactic burden of proving downward-closure or transitivity
side-conditions.
%
Additionally, \vett{} may serve as the basis of a future domain
specific proof assistant for category theoretic proofs.

@@

\item In \cref{sec:examples} we demonstrate how to use our syntax for
  formal category theory

\item In \cref{sec:semantics}, we develop some model theory for
  \vett{}, including a sound and complete notion of categorical model
  and sound interpretation in virtual equipments modeling ordinary,
  enriched and internal category theory.

@@

\begin{definition}[Virtual Equipment]
  A virtual equipment $\mathcal V$ consists of
  \begin{enumerate}
  \item A category $V_o$ of ``objects and vertical arrows''
  \item A chosen subset of the objects $V_s \subseteq V_o$ of ``small objects''
  \item A set ${\mathcal V}_h$ of ``horizontal arrows'' with source
    and target functions $s,t : {\mathcal V}_h \to {\mathcal V_o}^2$
  \item Sets of 2-cells of the following form:
    With appropriate ``multi-categorical'' notions of identity and
    composition.
\[\begin{tikzcd}% https://q.uiver.app/?q=WzAsNixbMCwwLCJDXzAiXSxbMCwyLCJEXzAiXSxbMiwwLCJDX3tufSJdLFsxLDAsIlxcY2RvdHMiXSxbMiwyLCJEXzEiXSxbMSwxLCJcXHBoaSJdLFszLDIsIlJfbiIsMCx7InN0eWxlIjp7ImJvZHkiOnsibmFtZSI6ImJhcnJlZCJ9fX1dLFswLDEsIkYiLDJdLFsyLDQsIkciXSxbMSw0LCJTIiwyLHsic3R5bGUiOnsiYm9keSI6eyJuYW1lIjoiYmFycmVkIn19fV0sWzAsMywiUl8wIiwwLHsic3R5bGUiOnsiYm9keSI6eyJuYW1lIjoiYmFycmVkIn19fV1d
	{C_0} & \cdots & {C_{n}} \\
	& \phi \\
	{D_0} && {D_1}
	\arrow["{R_n}", "\shortmid"{marking}, from=1-2, to=1-3]
	\arrow["f"', from=1-1, to=3-1]
	\arrow["g", from=1-3, to=3-3]
	\arrow["S"', "\shortmid"{marking}, from=3-1, to=3-3]
	\arrow["{R_0}", "\shortmid"{marking}, from=1-1, to=1-2]
\end{tikzcd}\]
    We say that the 2-cell $\phi$ has $S$ as codomain, the sequence
    $R_0 \ldots R_n$ as domain and call $f$ and $g$ the left and right
    ``frames'', or that $\phi$ is framed by $f$ and $g$.
  \end{enumerate}
  We say a virtual double category has \emph{split restrictions} when
  for any horizontal arrow $R : C \pto D$ and vertical arrows $f : C'
  \to C$ and $g : D' \to D$ there is a chosen horizontal arrow $R(f,g)
  : C' \pto D'$ with a cartesian 2-cell to $R$ framed by $f,g$ and
  these chosen cartesian lifts are functorial in $f,g$.

@@

%% %% Please use bibtex, 
\subsubsection*{Acknowledgements}
Grants stuff...? The authors would like to thank David Jaz Myers, Emily Riehl, Mike
Shulman, Dominic Verity for helpful feedback on this work.

\bibliographystyle{splncs04}
\bibliography{cats}
